---
title: "Financial Report"
author: "Robert Manni"
date: "10/9/2021"
output:
  html_document:
    theme:
      bootswatch: solar
    df_print: paged
  word_document: default

---

```{r, include=FALSE}
# This chunk contains the complete code used to generate the report for this Rmarkdown report excluding graphical analysis. 

library(tidyverse)
library(skimr)
library(scales)
library(bslib) # For html reports

# Details:

data = read.csv(choose.files(), na.strings = c(""))

# Inspect the data

# head(data) # Checks top few rows
# tail(data) # Checks the bottom few rows
# str(data) # checks the class of each variable (column)

#Or use skimr:
#skim(data)

data2.0 = data %>% 
  mutate(Expenses = gsub("Dollars", "", Expenses), # Removed "Dollars".
         Expenses = gsub(",", "", Expenses), # Removed commas and the space.
         Expenses = str_squish(Expenses),
         Expenses = as.numeric(Expenses), # Changed to integer (hopefully. Will need to check later.).
         Revenue = gsub("\\$", "", Revenue), # used esscape sequence '\\' to remove '$'. 
         Revenue = gsub(",", "", Revenue),# Removed commas and the space.
         Revenue = as.numeric(Revenue), # Makes Revenue MATHEMATICAL MAN!
         Profit = as.numeric(Profit), # Makes Profits MATHEMAICAL MAN!
         Growth = gsub("%", "", Growth), # Removed % symbol from values. 
         Growth = as.numeric(Growth)) %>% # Makes Growth MATHEMATICAL MAN!
  mutate(State = ifelse(City == 'San Francisco', 'CA', State), # NA value in State is CA
         State = ifelse(City == 'New York', 'NY', State), # NA value in State is NY
         Industry = ifelse(Name == 'Techline', 'IT Services', Industry), #NA value in Industry is IT Services
         Industry = ifelse(Name == 'Cityace', 'Retail', Industry), #NA value in Industry is Retail
         Expenses = ifelse(is.na(Expenses), Revenue-Profit, Expenses), # Calculated what Expenses are from existing columns. 
         Inception = ifelse(Name == 'Lathotline', '2011', Inception)) %>%  # Made the inception date 2011 for the NA value.
  drop_na() #drops NA's for the two construction columns that I cannot infer data for.

# Verify data is tidy:
#skim(data)




```

# Scope

This report is to elucidate from the data the trends between different industries and the relationship between growth, profit and overall revenue in these different industries from a dataset of 500 companies. 

## Deliverables:
1. A graph displaying revenue, expenses and profit by industry. 
2. A graphical representation of the relationship between expenses and revenue by industry. 
3. A graphical representation of growth by industry. 

# Results

## Overview:

```{r, echo=FALSE}
plot = ggplot(data = data2.0) # Assign data to plot to a veriable. I just like that. 

plot + geom_point(aes(x = Revenue, y = Expenses,
                      size = Profit,
                      color = Industry))  +
  labs(title = "Revenue by Expense Growth",
       subtitle = "By Insdusry",
       caption = "Profit is inversely related to Expenses.") +
  scale_x_continuous(NULL,
                     labels = label_dollar(),
                     guide = guide_axis(angle = 45)) +
  scale_y_continuous(NULL,
                     labels = label_dollar())
```

## Relationship of Revenue Growth to Expenses by Industry

```{r, echo=FALSE}
plot + geom_point(aes(x = Expenses, y = Revenue,
                      color = Industry)) +
  geom_smooth(aes(x = Expenses, y = Revenue,
                      color = Industry),
              fill = NA, size  = 1.2) + 
  labs(title = "Relationship of Revenue vs Expenses",
       caption = "Government Services seems pretty steady!") +
  scale_x_continuous(NULL,
                     breaks = breaks_extended(n = 7),
                     labels = scales::label_dollar(),
                     guide = guide_axis(angle = 45)) +
  scale_y_continuous(NULL,
                     breaks = breaks_extended(n = 7),
                     labels = scales::label_dollar(),)
  
```

## Overall Growth by Industry

```{r, echo=FALSE}
plot + geom_boxplot(aes(x = Industry, y = Growth, 
                        color = Industry)) +
  labs(title = "Growth by Industry",
       y = "Growth (%)") +
  scale_x_discrete(NULL,
                   guide = guide_axis(angle = 45))
```

