---
title: "Financial Report"
author: "Robert Manni"
date: "10/9/2021"
output:
  html_document:
    theme:
      bootswatch: solar
    df_print: paged
  word_document: default

---

```{r, include=FALSE}
# This chunk contains the complete code used to generate the report for this Rmarkdown report excluding graphical analysis. 

library(tidyverse)
library(skimr)
library(scales)
library(bslib) # For html reports
library(plotly)

# Details:

data = read.csv(choose.files(), na.strings = c(""))

# Inspect the data

# head(data) # Checks top few rows
# tail(data) # Checks the bottom few rows
# str(data) # checks the class of each variable (column)

#Or use skimr:
#skim(data)

data2.0 = data %>% 
      mutate(Expenses = as.numeric(gsub("[\\$, Dollars]", "", Expenses)),  # Remove '$', ',', and " Dollars"
             Revenue = as.numeric(gsub("[\\$,]", "", Revenue)),    # Remove '$' and ','
             Profit = as.numeric(Profit), 
             Growth = as.numeric(gsub("%", "", Growth))
  ) %>% 
  mutate(State = case_when(City == 'San Francisco' ~ 'CA', 
                           City == 'New York' ~ 'NY', 
                           TRUE ~ State),  # Keep original State if no match
         Industry = case_when(Name == 'Techline' ~ 'IT Services', 
                              Name == 'Cityace' ~ 'Retail', 
                              TRUE ~ Industry),  # Keep original Industry if no match
         Expenses = ifelse(is.na(Expenses), Revenue - Profit, Expenses),
         Inception = ifelse(Name == 'Lathotline', '2011', Inception)) %>% 
  drop_na()
#drops NA's for the two construction columns that I cannot infer data for.

# Verify data is tidy:
#skim(data)




```

# Scope

This report is to elucidate from the data the trends between different industries and the relationship between growth, profit and overall revenue in these different industries from a dataset of 500 companies. 

## Deliverables:
1. A graph displaying revenue, expenses and profit by industry. 
2. A graphical representation of the relationship between expenses and revenue by industry. 
3. A graphical representation of growth by industry. 

# Results

## Overview:

```{r, echo=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

plot = ggplot(data = data2.0) # Assign data to plot to a veriable. I just like that. 

Overview_plot = plot + geom_point(aes(x = Revenue, y = Expenses,
                      size = Profit,
                      color = Industry))  +
  labs(title = "Revenue by Expense Growth",
       subtitle = "By Insdusry",
       caption = "Profit is inversely related to Expenses.") +
  scale_x_continuous(NULL,
                     labels = label_dollar(),
                     guide = guide_axis(angle = 45)) +
  scale_y_continuous(NULL,
                     labels = label_dollar())

ggplotly(Overview_plot)
```

## Relationship of Revenue Growth to Expenses by Industry

```{r, echo=FALSE}
Rev_Growth = plot + geom_point(aes(x = Expenses, y = Revenue,
                      color = Industry)) +
  geom_smooth(aes(x = Expenses, y = Revenue,
                      color = Industry),
              fill = NA, size  = 1.2) + 
  labs(title = "Relationship of Revenue vs Expenses",
       caption = "Government Services seems pretty steady!") +
  scale_x_continuous(NULL,
                     breaks = breaks_extended(n = 7),
                     labels = scales::label_dollar(),
                     guide = guide_axis(angle = 45)) +
  scale_y_continuous(NULL,
                     breaks = breaks_extended(n = 7),
                     labels = scales::label_dollar(),)
 
ggplotly(Rev_Growth) 
```

## Overall Growth by Industry

```{r, echo=FALSE}
Overall_Report = plot + geom_boxplot(aes(x = Industry, y = Growth, 
                        color = Industry)) +
  labs(title = "Growth by Industry",
       y = "Growth (%)") +
  scale_x_discrete(NULL,
                   guide = guide_axis(angle = 45))

ggplotly(Overall_Report)
```

