---
title: "Global Maritime Pirate Attacks Project"
author: "Robert Manni"
date: "2024-02-13"
output: html_document
---

### Global Maritime Pirate Attacks (1993–2020)

Dataset and description provided by VAGIF of Kaggle. 

### Phase 1 | Exporatory Data Analysis

## Cleaning and Preparation of the Data

# Loading the Data

```{r}
#packages:
library(tidyverse) # tidyverse for data handling.
library(lubridate) # for working with dates and times
library(sf) # for working with spatial data
#library(rnaturalearth) # country specific data not sure if actually needed (evaluate)

#set working directory

data_pirates = read.csv(choose.files()) # choose.files is nice, and for some reason it's not grabbing the files in the working directory anyway.
data_countryCodes = read.csv(choose.files())
data_countryIndicators = read.csv(choose.files())

# glimpse data to see data types, NA values, and other high level information
glimpse(data_pirates)
glimpse(data_countryCodes)
glimpse(data_countryIndicators)

# NA values are present in much of the dataset - determining how much may be helpful for more specific observations. 

data_pirates %>% # the data is getting coerced, so might try another method
  select(attack_type) %>% # Number of NA values in attack type
  is.na() %>% # find the NA values
  sum() # how many we have in total: 120

# how do I want to handle/treat NA values in this instance? I'm thinking for analyzing attack by vessel and location, KNN would be a sufficient model as I don't believe this has a linear relationship and there will be value in classification of vessels.

# didn't see any missing "date" values. There are NA values in time - and are characters. So let's look into that. Most of the data has missing time values, so, using median, might be okay, but it's a huge chunk of missing data. 

# step 1: Because it is easier - is to map the data a graph. 

```

# Cleaning and Transformation

```{r}
# Pirate Attacks Data


  

```


## Data Exploration 

### Phase 2 | Data Vizualization 

### Phase 3 | ABM

##About Dataset

#Context
Maritime piracy and armed robbery against ship are one of the contemporary challenges of the maritime industry. These two phenomena have a global impact on maritime trade and security.
Nowadays, the Gulf of Aden and the Indian Ocean are considered high risk areas in terms of piracy and armed robbery against ships activities. In this regard, both the international community and the coastal States of the region have deployed every effort to try to find ways to address the problem.

#Content
This dataset contains information from more than 7,500 maritime pirate attacks that took place between January 1993 and December 2020, as well as country indicator data for the same time period. The pirate attack data was collected from the International Maritime Bureau (IMB), tidied, and augmented with geospatial data. The country indicator data was gathered from a variety of sources, notably The World Bank. The data is contained in Comma Separated Value (CSV) files.

pirate_attacks.csv
Date [Key] - Date of Attack. Used as a key with the Country Matrix data frame.
Time - Time the attack took place, either in UTC or Local Time.
Longitude - Longitude where the attack took place.
Latitude - Latitude where the attack took place.
Attack Type - Either NA (Missing), Attempted, Boarding, or Hijacked.
Location Description - A text description of the location. With attacks taking place at sea, it is not as simple as just naming a city or town.
Nearest Country [Key] - The country code whose shore is closest to the attack. The resolution is around 1 km, it can be much better depending on how detailed the mapping of the coast is in the vicinity.
EEZ Country [Key] - The Exclusive Economic Zone country code in which the attack took place, if it took place within an EEZ.
Shore Distance - Distance in kilometres to the shore from the attack location. This is the true geographic distance over the surface of the earth.
Shore Longitude - The longitude of the closest point on the shore to the attack.
Shore Latitude - The latitude of the closest point on the shore to the attack.
Attack Description - The text description of the attack if it exists.
Vessel Name - The name of the ship which was attacked if it is known.
Vessel Type - The type of vessel attacked if known.
Vessel Status - The status of the ship at the time it was attacked. Either NA (Missing), Berthed (Tied to a berth), Anchored (anchored at sea or in a harbour), or Steaming (ship underway).
country_indicators.csv
Country [Key] - The country in ISO3 country code format.
Corruption Index - Corruption Perceptions Index.
Homicide Rate - Total Intentional Homicides per 100,000 people.
GPD - Gross Domestic Product (US Dollars).
Total Fisheries Per Ton - Total Fisheries Production (metric tons).
Total Military - Total Number of Armed Forces personnel.
Population - Country Population.
Unemployment Rate - Percentage of the Country Unemployed.
Total GR - Total Government Revenue. An indication of how well the country collects taxes.
Industry - Industrial contribution to total GDP.
country_codes.csv
Country [Key] - The country in ISO3 country code format.
Region - The region the country is in.
Country Name - The English country name.
